Cover letter for the original IJRR submission

### 2017-10-02 reviews_tro/reviewer_response.tex

\newcommand\dx{\dot{x}}
\newcommand\dy{\dot{y}}
\newcommand\ddx{\ddot{x}}
\newcommand\ddy{\ddot{y}}
\def\real{{\mathbb R} }
\newcommand\quot[1]{\begin{quote} \underline{quote}: \textbf{#1}\end{quote}}
\newcommand\as[1]{\begin{quote} \underline{answer}: {#1}\end{quote} }
\newcommand\qt[1]{\begin{quote} \underline{added text in paper}: \textit{#1}\end{quote} \leavevmode \\ }
\newcommand\jp{ \leavevmode \\}
\frenchbsetup%
{%
StandardItemLabels=true,%
ItemLabels=\ding{43},%
}%
\DeclareUnicodeCharacter{00A0}{ }
\author {}
\title {Cover letter for the resubmission of our paper ``An efficient acyclic contact planner for multiped robots''}
\date {}
\begin{document}
\maketitle

\section{Answers to reviewer 1}
\ndone reecriture p3 + enlever main video

\quot {As for novel contributions, this did not seem to be addressed/explained
until page 3 (in the last paragraph of Sec. I). Much of the methodology
has already been presented in earlier work, and the particular aspects
that set THIS paper apart seem (as claimed by the authors) to be: 
}

\as{Thanks for the comment. Indeed, all reviewers agreed that the contribution with respect to previous work was not clearly stated. We updated the paper to reflect the following. Firstly, our planner is almost identical to the one in [20], with important additions that are the pseudo code for the algorithm (the algorithm is not given in [20]), as well as the introduction of a more efficient criterion
for checking static equilibrium (see next question). The optimal control is no longer a claimed contribution of the paper, even though it differs from [1] in its formulation. The main
novelty of the paper resides in the experimental setup and analysis, as well as the empirical proof that the planner works with real robots that are more constrained kinematically than the virtual avatars from [20]. We believe that this paper is thus in accordance with the guidelines of T-RO regarding the extension of a conference paper.}
\qt{The present paper is an extension of our conference paper [20]. As such, the motion planning algorithm presented here is the same as the one presented in the
previous work. Section IV and V of the present paper are thus paraphrasing the original paper, with the exception of Section V.B.2, where we additionally provide the pseudo-code of the algorithm absent from the original paper. We also improve the computational performance of the method by replacing the static equilibrium test with a more efficient test based on an new LP formulation presented in Section VI. \\
Compared with the original paper, we also release our source code, and provide the complete procedure to use our planner for any legged robot. We thus provide
all the technical information required to use the planner. \\
To validate our approach, in this paper we had to implement or integrate other tools based on previous publications from our team. We implemented a solution to $\mathcal{P}_3$ based on a reformulation of our previous work[1]. The implementation of this solution is also provided as an open-source project, which can be integrated directly in our framework. Our solution to $\mathcal{P}_3$ then allows us to verify that the synthesized motions are physically consistent, using our implementation of the state of the art simulator proposed in \textcolor{red}{TODO}.
We do not consider these additions as contributions per se, but describing them here is required to understand the experimental contribution of this paper.
Indeed, thanks to these tools we have been able to design a complete pipeline for synthesizing multi contact motions from scratch.
The novelty of the present paper lies in the rigorous experimental setup we propose to demonstrate its interest.
As opposed to our previous paper, we demonstrate the algorithms with actual robotic models. These models are much more constrained kinematically than the virtual avatars used before,
thus making the planning problem even harder. Through an extensive empirical analysis, we demonstrate that the method remains interesting in this context, qualitatively and efficiently.}
\done

@@

\as{The main contribution of the paper is the proposal and validation of an algorithm for P1 and P2. 
In the new version of the paper, we do not consider that our solution to P3
is a new contribution, although our formulation differs from [1] (our control variable is the acceleration
of the center of mass, whereas in [1] the control variable is the contact forces). We provide it here for completeness, because it corresponds to our implementation, and 
was required to validate our approach. 
We rephrased the text to clearly indicate that the algorithm is not a contribution. We made the choice not to give an intuition nor describe the solver in the core
of the paper because it is not required to understand the contributions to P1 and P2.}
\qt{ We implemented a solution to $\mathcal{P}_3$ based on a reformulation of our previous work[1]. The implementation of this solution is also provided as an open-source project, which can be integrated directly in our framework. Our solution to $\mathcal{P}_3$ then allows us to verify that the synthesized motions are physically consistent, using our implementation of the state of the art simulator proposed in \textcolor{red}{TODO}.
We do not consider these additions as contributions per se, but describing them here is required to understand the experimental contribution of this paper.}\done

\quot{
(4)	The last contribution is validation using an in-house
simulator, as opposed to more approximate (unrealistic avatar) models
used in previous work.	As with the previous contributions claimed, you
should flesh out just a bit more detail, along with a roadmap to where
a particular contribution appears in the paper, to prepare and guide
the reader a bit more. More precisely, what are the differences between
these two simulation approaches (previous and current)? You need a
better description of what your in-house simulator does and does not
do; a vague citation to another paper based on that simulator [38] is
not sufficient.}

\as{Thanks for the comment. Regarding the absence of details on the dynamic simulator, first of all we apologize because [38] is not the appropriate reference.
We added a section VIII.A where we describe the simulation in more details.

Regarding the difference between our previous and current work, the postures computed in [20] were never validated in a simulator, nor the complete motions since 
we did not have a solution to P3 at the time. Then the main
difference between the avatars of [20] and the real robots we now address are the joint limitations and auto-collisions probabilities.
The robots are much more constrained in this regard, which makes the planning harder. We thus had to demonstrate that our algorithm is also working
for these models, while maintaining good performance.
We added this missing piece of information to the paper, at the end of Section I.B.}
\qt{(Section VIII.A) To generate continuous movements from our contact plans we used either the framework proposed in [1], or our own implementation of a $\mathcal{P}_3$ solver (Appendix C).
The resulting movements have been validated either on the real HRP-2 robot (details can be found in [1]), or with our dynamic simulator, based on a state-of-the-art algorithm [25].
In the simulations we controlled the robot with a standard inverse-dynamics controller[31].
This controller tries to follow the given whole-body trajectories, giving higher priority to the center-of-mass and end-effectors tracking with respect to the joint tracking.
The controller also makes sure that the resulting contact forces lie inside the specified friction cones (we used a friction coefficient of 0.3), and that the joint position, velocity and torque limits are satisfied.
The companion video shows the obtained motions.}
\done
%~ \textcolor{red}{TODO ANDREA}.

@@

\as{Thanks for the comment. The choice of not detailing too much the environment and octree description goes along with 
our effort to maintain the paper as readable as possible, since it already contains various concepts. However,
we agree that additional information must be added. In Section III we now indicate that the environment is represented as a polygon soup (or 3D mesh), and
that we assume that it is fully known. Although the integration of on-line sensing and state estimation in the planner is an issue currently considered by our team, it is indeed
out of the scope of this paper. We thus also indicate that we assume the environment and state of the robot are fully known. To perform the intersection computation, we use
the fcl collision detection library, which natively proposes the intersection method. We added a reference to the library in the paper. The performance of the method is included
in the ``collision'' column of Table II. We updated the legend of Table II to indicate that this time is negligible in comparison to the other collision detections that occur.
} 
\qt{(Section III) \textbf{The environment} $O$ is defined as the union of the obstacles $O_i$ that it contains. $O$ is represented
as a polygon soup (or mesh), where the normal of each surface is known. No further requirement is needed
by our approach. In this work, we assume the environment is fully known. State uncertainty is out of the scope of the paper. \\
(Legend of Table II)  The Collision column times includes the (negligible) octree intersection operation necessary to retrieve the candidate samples.}
\done

\quot{I appreciate the thorough presentation of results in the paper, however
it seems that this presentation is somewhat biased towards simulation
results. Simulations are perfectly OK for planning algorithms, but if
you decided to present the experimental results (HRP-2), then at least some
illustrations should be included in the paper. now they are only in the
accompanying video.  also a comparison between the results of the same
task(s) executed in dynamic simulation and on the real robot should
convince the reader that the planning method's good performance does
not depend too much on the idealized simulation environment. }
\as{We agree with the remark, and included some illustrations of the experimental results on HRP-2 in the paper. \\
We now elaborate on the validity of the simulation with respect to an experiment on the real robot. As also suggested by reviewer 1, we added a paragraph
in the paper (Section VIII.A)  to provide more information on the simulation we use to validate our motions. This simulation is an implementation of the state of the art approaches used by the community. Of course, it does not guarantee that the motion could be executed on the robot, but it is used extensively by the community as a reasonable feasibility test. The suggestion of the reviewer to compare the simulation results with those on the real robot is interesting. However, our robot HRP-2 is not torque controlled, but position controlled, therefore as of today a comparison is not possible.}
\qt{(Section VIII.A) To generate continuous movements from our contact plans we used either the framework proposed in [1], or our own implementation of a $\mathcal{P}_3$ solver (Appendix C).
The resulting movements have been validated either on the real HRP-2 robot (details can be found in [1]), or with our dynamic simulator, based on a state-of-the-art algorithm [25].
In the simulations we controlled the robot with a standard inverse-dynamics controller[31].
This controller tries to follow the given whole-body trajectories, giving higher priority to the center-of-mass and end-effectors tracking with respect to the joint tracking.
The controller also makes sure that the resulting contact forces lie inside the specified friction cones (we used a friction coefficient of 0.3), and that the joint position, velocity and torque limits are satisfied.
The companion video shows the obtained motions.}
\done

\section{Answers to reviewer 6}
\quot{ My primary concern, however,
is that the elements described in this paper have been described in
previous published work, which makes the particular contributions of
this work unclear. To be specific, the reachability-guided planner
appears identical to the planner in [20] (as stated by the authors in
the end of Section I.B.), and the optimal control formulation is from [1]
(again as stated in I.B.).}
\as{Thanks for the comment. Indeed, all reviewers agreed that the contribution with respect to previous work was not clearly stated. We updated the paper to reflect the following. Firstly, our planner is almost identical to the one in [20], with important additions that are the pseudo code for the algorithm (the algorithm is not given in [20]), as well as the introduction of a more efficient criterion
for checking static equilibrium (see next question). The optimal control is no longer a claimed contribution of the paper, even though it differs from [1] in its formulation. The main
novelty of the paper resides in the experimental setup and analysis, as well as the empirical proof that the planner works with real robots that are more constrained kinematically than the virtual avatars from [20]. We believe that this paper is thus in accordance with the guidelines of T-RO regarding the extension of a conference paper.}
\qt{(Section I.B) The present paper is an extension of our conference paper [20]. As such, the motion planning algorithm presented here is the same as the one presented in the
previous work. Section IV and V of the present paper are thus paraphrasing the original paper, with the exception of Section V.B.2, where we additionally provide the pseudo-code of the algorithm absent from the original paper. We also improve the computational performance of the method by replacing the static equilibrium test with a more efficient test based on an new LP formulation presented in Section VI. \\
Compared with the original paper, we also release our source code, and provide the complete procedure to use our planner for any legged robot. We thus provide
all the technical information required to use the planner. \\
To validate our approach, in this paper we had to implement or integrate other tools based on previous publications from our team. We implemented a solution to $\mathcal{P}_3$ based on a reformulation of our previous work[1]. The implementation of this solution is also provided as an open-source project, which can be integrated directly in our framework. Our solution to $\mathcal{P}_3$ then allows us to verify that the synthesized motions are physically consistent, using our implementation of the state of the art simulator proposed in \textcolor{red}{TODO}.
We do not consider these additions as contributions per se, but describing them here is required to understand the experimental contribution of this paper.
Indeed, thanks to these tools we have been able to design a complete pipeline for synthesizing multi contact motions from scratch.
The novelty of the present paper lies in the rigorous experimental setup we propose to demonstrate its interest.
As opposed to our previous paper, we demonstrate the algorithms with actual robotic models. These models are much more constrained kinematically than the virtual avatars used before,
thus making the planning problem even harder. Through an extensive empirical analysis, we demonstrate that the method remains interesting in this context, qualitatively and efficiently.}\done

\quot{The authors highlight a particular
contribution of this work to be the efficient robust balance criterion,
but I had trouble understanding this particular contribution. Does this
refer to the approximation of equilibrium feasibility by contact
reachability? In that case, how does this differ from just applying the
planner in [20] directly? Or does this refer to the robustness
parameter $b_0$ in Appendix B Eq. (15)?}
\as{Indeed, this part was also not clear. The robust balance criterion (renamed equilibrium criterion) indeed refers to the parameter $b_0$ in Appendix B (now Section VI). The solved
LP replaces the static equilibrium test performed in [20], which was not computationally efficient, and took up to 4ms per candidate to compute. We updated the paper to explain
this in details.}
\qt{(Section I.B) We also improve the computational performance
of the method by replacing the static equilibrium test with a more efficient test based on an
new LP formulation presented in Section VI.}
\done

\quot{In addition, sections IV and V are very similar to the description of
the planner in [20]. That's not surprising, of course, but there is no
indication in those sections that the work being presented is a
rephrasing of an earlier publication.[\dots] }
\as{We agree with the comment, and indeed made this clearer in the new submission.}
\qt{(Section I.B) Section IV and V of the present paper are thus paraphrasing the original paper, with the exception of Section V.B.2, where we additionally provide the pseudo-code of the algorithm absent from the original paper.}\done

\quot{I do, however,
believe that the fact that the technical approaches are rephrasings of
prior work should be made more clear (or that sections which can be
replaced by citations of [20] and [1] be removed).}
\as{Thanks for the advice. This suggestion was in contradiction with the other reviewers, who thought that the paper should be self-contained.
The associate editor decided that the paper must be self contained, so we maintained the indicated sections, but made clearer the fact that they are rephrasing of previous work. }\done

\quot{*I think I understand what the authors mean by solving at
"interactive" rates, but this is worth defining. In my mind, it means
something like "before the human operator gets bored", or roughly > 0.5
Hz. But since the authors make the claim that this is the "first
interactive implementation of a contact planner", then that statement
should be made quantitative and verifiable}
\as{We agree with the relevant comment, and updated the paper accordingly.}
\qt{(End of Section I.A.) We define an interactive planner as one that outputs a solution in a few seconds or less.}\done

\quot{* In Section I.A. the authors refer to Bretl's work on multi-contact
planning as the inspiration for the P1,P2,P3 decomposition of the
problem. But it is not, as the authors state, always necessary to
decompose motion planning into choosing a series of quasi-static
configurations. There are a variety of feasible motions with no static
equilibrium configurations, so it is worth mentioning that this
decomposition only solves a subset of motion planning problem.}
\as{We apologize if this was not made clear enough in our first submission. In the conclusion, we indicate that the planner is limited to specific scenarios (where quasi-static motion
is always an option) ``\textit{Our method applies to any scenario where at least one
contact friction cone contains the direction opposed to the
gravity (i.e. quasi-flat)[\dots]}''. We updated the state of the art to indicate that the decomposition in general is restricitve.}
\qt{(End of before last paragraph of Section I.A) A limitation with
this kind of approach, shared by our own method, is that most of the existing planners only address a subset of the problem, because their ability to find a solution depends
on the existence of quasi-static equilibrium configurations along a feasible path, which is too restrictive in the general case.}\done

### 2017-10-02 body/conclusion.tex
sota update on bretls work

\subsection{Accuracy and performance gain of the reachability condition}

@@

\subsection{Addressing $\mathcal{P}_3$}
\adnote{This subsection seems redundant to me}
This paper focuses on contact planning of multi-contact locomotion.It does not address directly the continuous motion generation
. Problem $\mathcal{P}_3$ remains an active research topic~\cite{Carpentier2016, herzog2015trajectory},
for which no off-the shelf solution exists thus far. We thus implemented a dedicated solution, described in Appendix~\ref{app:optim}.
The companion video shows the results that we obtained in simulation and on the real HRP-2 robot.
In these scenarios the robot is able to execute our plans, while respecting the constraints inherent to the problem: joint kinematics and torque limits and
collision avoidance.
%~ \adnote{Probably you should add in appendix an explanation of how you solve $P_3$ and refer to it here.}

@@

%~ \subsection{feasibility }

@@

Our method applies to any scenario where at least one contact friction cone contains
the direction opposed to the gravity (i.e. \textit{quasi-flat}). This class of scenarios include all the problems proposed at the DARPA Robotics Challenge.
One way to further extend its range of application, which we consider for future work, is to include the equilibrium criterion when solving $\mathcal{P}_1$.
Considering the set of obstacles intersecting with the reachable workspace for a given root configuration as candidate surfaces, we can use them to verify the equilibrium criterion.
This would give us a necessary condition for \equilibriumfeasibility.

>>>>

%~ \subsection{Accuracy and performance gain of the reachability condition}

@@

%~ \subsection{Addressing $\mathcal{P}_3$}
%~ \adnote{This subsection seems redundant to me}
%~ This paper focuses on contact planning of multi-contact locomotion.It does not address directly the continuous motion generation
%~ . Problem $\mathcal{P}_3$ remains an active research topic~\cite{Carpentier2016, herzog2015trajectory},
%~ for which no off-the shelf solution exists thus far. We thus implemented a dedicated solution, described in Appendix~\ref{app:optim}.
%~ The companion video shows the results that we obtained in simulation and on the real HRP-2 robot.
%~ In these scenarios the robot is able to execute our plans, while respecting the constraints inherent to the problem: joint kinematics and torque limits and
%~ collision avoidance.

@@

@@

Our method applies to any scenario where at least one contact friction cone contains
the direction opposed to the gravity (i.e. \textit{quasi-flat}). This class of scenarios include all the problems proposed at the DARPA Robotics Challenge.
One way to further extend its range of application, which we consider for future work, is to include the equilibrium criterion when solving $\mathcal{P}_1$.
Considering the set of obstacles intersecting with the reachable workspace for a given root configuration as candidate surfaces, we can use them to verify the equilibrium criterion.
This would give us a necessary condition for \equilibriumfeasibility. %, which we investigate in~\cite{fernbach:hal-01486933}.

### 2017-10-02 body/intro.tex
sota update on bretls work

\textbf{On the other hand, the two problems \Pa and \Pb might be decoupled} to reduce the complexity. The feasibility and interest of the decoupling has been shown by Escande et al. \cite{DBLP:conf/iser/EscandeKMG08} who manually set up a rough root guide path (i.e. an ad-hoc solution to $\mathcal{P}_1$), and then addressed \Pb as the combinatorial computation of a feasible contact sequence in the neighborhood of the guide. A solution could then be found, %efficiently when considering quadruped locomotion \citep{kalakrishnan2011learning}, 
but at the cost of prohibitive computation times (up to several hours) for constraining scenarios. This approach is suboptimal because the quality of the motion depends on the quality of the guide path. Bouyarmane et al. ~\cite{Bouyarmane2009} precisely focused on automatically computing a guide path with guarantees of \textit{equilibrium feasibility}, by extending key frames of the path into whole-body configurations in static equilibrium. Randomly-sampled configurations are projected to the contact manifold using an inverse-kinematics solver, a computationally-expensive process (about 15 minutes to compute a guide path in the examples presented). Moreover this explicit projection is insufficient to guarantee the feasibility between two key postures in the path. Chung and Khatib~\cite{7140082} also proposed a decoupled approach, with a planning phase based on the reachable workspace of the robot limbs, used to judge the ability to make contact with a discretized environment. This planning phase does not account for collisions, implying that re-planning is required in case of failure. In highly-constraining cases such as the car egress scenario we address, we believe that including collision constraints in the planning is a requirement~\citep{tonneauisrr15,grey2017footstep}.

@@

As far as robotics applications are concerned, none of the existing planners is \gls{interactive}.
However, recent contributions to the interpolation between contact poses (problem $\mathcal{P}_3$) have brought promising preliminary solutions \citep{Hauser2014, herzog2015trajectory, Park116, Carpentier2016}. In particular, our algorithm proposed in \cite{Carpentier2016} is \gls{interactive}.
Therefore, a planner capable of efficiently solving \Pa and \Pb could outperform all existing planners if coupled with an interpolation method solving $\mathcal{P}_3$.
The main contribution of this paper is exactly this planner.

>>>>

\textbf{On the other hand, the two problems \Pa and \Pb might be decoupled} to reduce the complexity. The feasibility and interest of the decoupling has been shown by Escande et al. \cite{DBLP:conf/iser/EscandeKMG08} who manually set up a rough root guide path (i.e. an ad-hoc solution to $\mathcal{P}_1$), and then addressed \Pb as the combinatorial computation of a feasible contact sequence in the neighborhood of the guide. A solution could then be found, %efficiently when considering quadruped locomotion \citep{kalakrishnan2011learning}, 
but at the cost of prohibitive computation times (up to several hours) for constraining scenarios. This approach is suboptimal because the quality of the motion depends on the quality of the guide path. Bouyarmane et al. ~\cite{Bouyarmane2009} precisely focused on automatically computing a guide path with guarantees of \textit{equilibrium feasibility}, by extending key frames of the path into whole-body configurations in static equilibrium. Randomly-sampled configurations are projected to the contact manifold using an inverse-kinematics solver, a computationally-expensive process (about 15 minutes to compute a guide path in the examples presented). Moreover this explicit projection is insufficient to guarantee the feasibility between two key postures in the path. Chung and Khatib~\cite{7140082} also proposed a decoupled approach, with a planning phase based on the reachable workspace of the robot limbs, used to judge the ability to make contact with a discretized environment. This planning phase does not account for collisions, implying that re-planning is required in case of failure. In highly-constraining cases such as the car egress scenario we address, we believe that including collision constraints in the planning is a requirement~\citep{tonneauisrr15,grey2017footstep}. A limitation with
these approaches (including our method) is that the existing planners only address a subset of the problem, because their ability to find a solution depends
on the existence of quasi-static equilibrium configurations along a feasible path, which is too restrictive in the general case.

@@

As far as robotics applications are concerned, none of the existing planners is \gls{interactive}(We define an interactive planner as one that outputs a solution in a few seconds or less) . That is 
However, recent contributions to the interpolation between contact poses (problem $\mathcal{P}_3$) have brought promising preliminary solutions \citep{Hauser2014, herzog2015trajectory, Park116, Carpentier2016}. In particular, our algorithm proposed in \cite{Carpentier2016} is \gls{interactive}.
Therefore, a planner capable of efficiently solving \Pa and \Pb could outperform all existing planners if coupled with an interpolation method solving $\mathcal{P}_3$.
The main contribution of this paper is exactly this planner.
